
Reproducible Research: Peer Assessment 1
========================================================
This is a markdown document that contains the explanation and R code for my peer assesment 1 assignment for the
Coursera reproducible research course in the Data Science specilization.


## Loading and preprocessing the data

```{r, echo=TRUE}
activity <- read.csv("./activity.csv", header=TRUE)
head(activity)
tail(activity)
str(activity)
```

## What is mean total number of steps taken per day?

```{r, echo=TRUE}
library(ggplot2)
require(data.table)
activity_dt <- data.table(activity)

steps_by_day <- activity_dt[,data.frame(DailySteps= sum(steps,na.rm= TRUE)), by=date]
head(steps_by_day)

#plot(x= steps_by_day$date, y= steps_by_day$DailySteps, type="h")

plot(steps_by_day$DailySteps, type="h")


#calculate Mean
#mean_steps_by_day <- activity_dt[,data.frame(MeanDailySteps = mean(steps,na.rm= TRUE)), by=date]
#mean_steps_by_day

mean(steps_by_day$DailySteps)


#calculate Median
#median_steps_by_day <- activity_dt[,data.frame(MedianDailySteps = as.double(median(steps,na.rm= TRUE))), by=date]
#median_steps_by_day
median(steps_by_day$DailySteps)

```

## What is the average daily activity pattern?

```{r, echo=TRUE}
#calculate Mean steps for 5 minute interval
#mean_steps_by_day_5int <- activity_dt[activity_dt$interval==5,data.frame(MeanDailySteps5int = mean(steps,na.rm= TRUE)), by=date]

mean_steps_by_day_5int <- activity_dt[,data.frame(MeanDailySteps5int = mean(steps,na.rm= TRUE)), by="date,interval"]

head(mean_steps_by_day_5int)

#plot(x = mean_steps_by_day_5int$date, y= mean_steps_by_day_5int$MeanDailySteps5int, type= "l" )

#plot(mean_steps_by_day_5int$MeanDailySteps5int, type= "l" )

```

```{r fig.width=7, fig.height=6, echo=TRUE}
library(ggplot2)
#set.panel(2,1)
#qplot(x= mean_steps_by_day_5int_weekdays$interval, y=mean_steps_by_day_5int_weekdays$MeanDailySteps5intWday, group = mean_steps_by_day_5int_weekdays$typeOfDay, type = "l")

#xyplot(MeanDailySteps5intWday ~ interval | typeOfDay ,mean_steps_by_day_5int_weekdays, type="l", layout(1,2))

xyplot(MeanDailySteps5int ~ interval , mean_steps_by_day_5int, type="l")

#plot(mean_steps_by_day_5int$MeanDailySteps5int, type= "l" )

```



The 5-minute interval with the maximum number of steps in 5-minute interval 10

## Imputing missing values

```{r, echo=TRUE}

noNaVals <- complete.cases(activity)
activity_na <- activity[!noNaVals,]

#number of rows with NA values
nrow(activity_na)

#replace Na Values

activity_noNa <- activity

avgSteps <- mean(activity_noNa$steps, na.rm=TRUE)

random.impute <- function(a,b){
        naVal <- is.na(a)
        #n.naVal <- sum(naVal)
        #a.obs <- a[!naVal]
        imputed <- a
        imputeWith <- b
        imputed[naVal] <- imputeWith
        return(imputed)
}

activity_noNa2 <- random.impute(activity_noNa, avgSteps)

head(activity_noNa2)


require(data.table)
activity_noNa2 <- data.table(activity_noNa2)

steps_by_day_noNa <- activity_noNa2[,data.frame(DailySteps= sum(steps,na.rm= TRUE)), by=date]
head(steps_by_day_noNa)

plot(steps_by_day_noNa$DailySteps, type="h")

#calculate Mean
#mean_steps_by_day_noNa <- activity_noNa2[,data.frame(MeanDailySteps = mean(steps,na.rm= TRUE)), by=date]
#head(mean_steps_by_day_noNa)
mean(steps_by_day_noNa$DailySteps)

#calculate Median
#median_steps_by_day_noNa <- activity_noNa2[,data.frame(MedianDailySteps = as.double(median(steps,na.rm= TRUE))), by=date]
#head(median_steps_by_day_noNa)
median(steps_by_day_noNa$DailySteps)


```
There is a difference between the mean and median that was previously calculated when compared to the mean and median that was calculate with the imputed data.

The impact of imputing data on the estimates for total daily numbers is the date 2012-10-01 no longer has Zero daily steps.


## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}

activity_noNa2[,"dayOfWeek"] <- weekdays(as.Date(activity_noNa2$date))
activity_noNa2$dayOfWeek <- as.factor(activity_noNa2$dayOfWeek)
head(activity_noNa2)

activity_noNa2[,"typeOfDay"] <- ifelse(activity_noNa2$dayOfWeek %in% c("Saturday","Sunday"), "Weekend", "Weekday")

#require(data.table)


mean_steps_by_day_5int_weekdays <- activity_noNa2[,data.frame(MeanDailySteps5intWday = mean(steps,na.rm= TRUE)), by="interval,typeOfDay"]

head(mean_steps_by_day_5int_weekdays)


meanData_5int <- activity_noNa2[,mean(steps), by="interval,typeOfDay"]

head(meanData_5int)

```
Panel plot

```{r fig.width=7, fig.height=6, echo=TRUE}
library(ggplot2)
#set.panel(2,1)
#qplot(x= mean_steps_by_day_5int_weekdays$interval, y=mean_steps_by_day_5int_weekdays$MeanDailySteps5intWday, group = mean_steps_by_day_5int_weekdays$typeOfDay, type = "l")

xyplot(MeanDailySteps5intWday ~ interval | typeOfDay ,mean_steps_by_day_5int_weekdays, type="l", layout(1,2))

```